---
title: "R Notebook: C. sordidus growth analyses"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#Step 0: Load packages
```{r}
library(tidyverse)
library(brms)
library(tidybayes)
library(modelr) #for data_grid function
library(bayesplot)
library(gridExtra) #for plots
library(RColorBrewer) #for plots
library(cowplot) #for plots
library(scales) #for plots
```


#Step 1: Load data
```{r}

load("cso_data.Rdata")

##want to use lagoon, female, stage 3 only file:
head(cso_dat_lfm)
str(cso_dat_lfm)

###need to run log-log model to make it fit better
cso_dat_lfm$llength<-log(cso_dat_lfm$fork_length_cm)

#create indicator variable
cso_dat_lfm$tin<-ifelse(cso_dat_lfm$treatment=="ratty", 0, 1)


##re-level to make ratty the reference level, rather than birdy:
cso_dat_lfm$treatment_b<-relevel(cso_dat_lfm$treatment, "ratty")

##checking ranges
range(cso_dat_lfm$fork_length_cm)
range(cso_dat_lfm[cso_dat_lfm$treatment=="ratty",]$fork_length_cm) #15.7  25.0
range(cso_dat_lfm[cso_dat_lfm$treatment=="birdy",]$fork_length_cm) #15.2 to 26.8
range(cso_dat_lfm[cso_dat_lfm$treatment=="ratty",]$age_final) #1 to 6
range(cso_dat_lfm[cso_dat_lfm$treatment=="birdy",]$age_final) #1 to 8


#check sample sizes:
cso_dat_lfm%>%
 group_by(island)%>%
  tally(age_final)

##need to drop levels:
cso_dat_lfm<-
  cso_dat_lfm%>%
  droplevels()
levels(cso_dat_lfm$island)


```


##response to reviewers: compare VBGF models where allow K, both K & Linf, neither to vary:
```{r}
#need to specify priors for nl models
#based on priors in Graham et al. 2018 (who used same formula)
vbgf_prior_k_only<- c(
  prior(uniform(15.2, 53.6), nlpar = "Linf", lb = 15.2, ub = 53.6), #constraind between minimum size and twice maximum
  prior(uniform(0,2), nlpar = "K", lb=0, ub = 2), #constrained because must be > 0, and k not > 2
  prior(uniform(0, 15.2), nlpar = "L0", lb = 0, ub = 15.2), ###L0 constrained bewettn 0 and minimum size observed
  prior(normal(0, 1), nlpar = "K1") ###weak, to allow either positive or negative (or 0) birdy effect
)

#run model - K only----------------
vbgf_mod_k_only <- 
  brm(
    bf(
      llength ~ log(Linf-(Linf-L0)*exp(-(K+K1*tin)*age_final)),
      Linf~1, K~1, L0~1, K1~1,
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = vbgf_prior,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "vbgf_mod_k_only"
  )
print(vbgf_mod_k_only) 
plot(vbgf_mod_k_only, ask = FALSE)
pp_check(vbgf_mod_k_only)


#run model - K and Linf----------------
vbgf_prior_both<- c(
  prior(uniform(15.2, 53.6), nlpar = "Linf", lb = 15.2, ub = 53.6), #constraind between minimum size and twice maximum
  prior(uniform(0,2), nlpar = "K", lb=0, ub = 2), #constrained because must be > 0, and k not > 2
  prior(uniform(0, 15.2), nlpar = "L0", lb = 0, ub = 15.2), ###L0 constrained bewettn 0 and minimum size observed
  prior(normal(0, 1), nlpar = "K1"), ###weak, to allow either positive or negative (or 0) birdy effect
  prior(normal(0, 10), nlpar = "Linf1") ###weak, to allow either positive or negative (or 0) birdy effect
)

vbgf_mod_both <- 
  brm(
    bf(
      llength ~ log((Linf+Linf1*tin)-((Linf+Linf1*tin)-L0)*exp(-(K+K1*tin)*age_final)),
      Linf~1, K~1, L0~1, K1~1, Linf1~1,
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = vbgf_prior_both,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "vbgf_mod_both"
  )
print(vbgf_mod_both) 
plot(vbgf_mod_both, ask = FALSE)
pp_check(vbgf_mod_both)
##k trend is still positive


##run model - neither vary----------------
vbgf_prior_none<- c(
  prior(uniform(15.2, 53.6), nlpar = "Linf", lb = 15.2, ub = 53.6), #constraind between minimum size and twice maximum
  prior(uniform(0,2), nlpar = "K", lb=0, ub = 2), #constrained because must be > 0, and k not > 2
  prior(uniform(0, 15.2), nlpar = "L0", lb = 0, ub = 15.2) ###L0 constrained bewettn 0 and minimum size observed
)


vbgf_mod_none <- 
  brm(
    bf(
      llength ~ log(Linf-(Linf-L0)*exp(-(K)*age_final)),
      Linf~1, K~1, L0~1,
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = vbgf_prior_none,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "vbgf_mod_none"
  )
print(vbgf_mod_none) 
plot(vbgf_mod_none, ask = FALSE)
pp_check(vbgf_mod_none)


###model comparisons----------
loo_both<-loo(vbgf_mod_both, moment_match=TRUE)
loo_k<-loo(vbgf_mod_k_only, moment_match=TRUE)
loo_none<-loo(vbgf_mod_none, moment_match=TRUE)

loocomp<-loo_compare(loo_both, loo_k, loo_none)

print(loocomp, simplify = F)
#model with only k is the best, model with none varying is definitely the worst.


#look at model weights: 
model_weights(vbgf_mod_both, vbgf_mod_k_only, vbgf_mod_none, 
              weights = "waic") %>% 
  round(digits = 2)
###matches results above: most weight for k only model (0.66) versus both model (0.31) versus none model (0.03)

##don't just blindly go on model weights - see what parameter estimates look like as well

###compare estimates:                 Estimate  Est.Error l-95% CI u-95% CI                 Estimate  Est.Error l-95% CI u-95% CI
print(vbgf_mod_k_only)#K1_Intercept       0.10      0.05     0.03     0.22 #K_Intercept        0.28      0.07     0.17     0.44
print(vbgf_mod_both)  #K1_Intercept       0.09      0.12    -0.16     0.32 #K_Intercept        0.27      0.12     0.10     0.55 

hypothesis(vbgf_mod_k_only, "K1_Intercept > 0") #posterior probability = 1
hypothesis(vbgf_mod_both, c("K1_Intercept > 0", #posterior probability = 0.78 - still quite strong evidence for an effect
                            "Linf1_Intercept < 0")) #posterior probability = 0.49 - absolutely no evidence that Linf differs by bird versus rat status

##Use model with only k varying because:
#(1) matches Graham et al. 2018 so facilitates comparisons
#(2) allows focus on variable of interest (growth rate)
#(3) most evidence for this model (from WAIC)
#(4) no evidence that Linf differs by bird-rat status (from hypothesis testing with both model)



#Compile posterior distributions:
vbgf_post <- posterior_samples(vbgf_mod_k_only)
median_hdi(vbgf_post$b_K1_Intercept, .width = .95) ## 0.09591383	0.01887863	0.1972184	
median_hdi(vbgf_post$b_K_Intercept, .width = .95) ##0.2747117	0.1553863	0.4144186	
median_hdi(vbgf_post$b_Linf_Intercept, .width = .95) ##24.62626	22.87138	26.89462	


vbgf_post_linf <- posterior_samples(vbgf_mod_both)
median_hdi(vbgf_post_linf$b_K1_Intercept, .width = .95) ## 0.08986419	-0.1668529	0.3058683	
median_hdi(vbgf_post_linf$b_Linf1_Intercept, .width = .95) ## 0.05632566	-7.510531	4.834921
median_hdi(vbgf_post_linf$b_Linf_Intercept, .width = .95) ## 24.86945	21.38321	32.38188	0.95	median	hdi
                                                          ##24.86945	32.65962	32.65962	0.95	median	hdi
median_hdi(vbgf_post_linf$b_K_Intercept, .width = .95) ## 0.258129	0.072564	0.4999814	
print(vbgf_mod_both)


```


#Final VBGF model (also note earlier versions built up models slowly, compared different priors, etc. - all came to same conclusions)
```{r}
#double-check diagnostics, compile posterior distributions, etc. for k-only model from above:

print(vbgf_mod_k_only) 
plot(vbgf_mod_k_only, ask = FALSE)
pp_check(vbgf_mod_k_only)
pairs(vbgf_mod_k_only)
prior_summary(vbgf_mod_k_only)
##all looks good

#compare posterior vs. prior distributions
#Linf:
posterior_samples(vbgf_mod_k_only) %>% 
  select(b_Linf_Intercept, prior_b_Linf) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#K:
posterior_samples(vbgf_mod_k_only) %>% 
  select(b_K_Intercept, prior_b_K) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope (rat):
posterior_samples(vbgf_mod_k_only) %>% 
  select(b_L0_Intercept, prior_b_L0) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope (rat):
posterior_samples(vbgf_mod_k_only) %>% 
  select(b_K1_Intercept, prior_b_K1) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()


#Compile posterior distributions:
vbgf_post <- posterior_samples(vbgf_mod_k_only)
median_hdi(vbgf_post$b_K1_Intercept, .width = .95) #median = 0.09591383, min = 0.01887863, max = 0.1972184
median_hdi(vbgf_post$b_K_Intercept, .width = .95) #ratty: 0.2747117
median_hdi(vbgf_post$b_K_Intercept+vbgf_post$b_K1_Intercept, .width = .95) #birdy: median = 0.3752644


#test hypothesis that birdy effect is > 0
hypothesis(vbgf_mod_k_only, "K1_Intercept  > 0") ##evidence ratio = 1332.33, posterior prob = 1

 
##calculate percent difference:
 vbgf_mod_k_only %>%
  spread_draws(b_K1_Intercept, b_K_Intercept) %>%
   median_qi(k_rat = b_K_Intercept, 
            k_bird_diff = b_K1_Intercept)%>%
   mutate(percent_diff = (k_bird_diff)/((k_rat+k_rat+k_bird_diff)/2)*100, #29.72518
          percent_change_rat_to_bird = (k_bird_diff)/k_rat*100) #34.91435


```


###posterior plots and plots with raw data and regression lines for VBGF model
```{r}

##posterior:
vbgf_k_post_plot <-
vbgf_mod_k_only %>%
  spread_draws(b_K1_Intercept) %>%
  ggplot(aes(y = 0, x = b_K1_Intercept, fill = stat(x < 0))) +
  stat_halfeyeh(point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) +
  scale_fill_manual(values = alpha(c("#005AB5", "#DC3220"),0.3))+
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() + 
  xlab("Difference in growth paramater k")+ 
  ylab("Posterior density") +
  scale_y_continuous(breaks = NULL) +
#  scale_x_continuous(limits=c(-0.1, 0.4), breaks=seq(-0.1, 0.3, .1), labels = scales::number_format(accuracy = 0.1))+
  #scale_x_continuous(limits=c(-0.05, 0.25), breaks=seq(0, 0.25,.1))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
vbgf_k_post_plot


##raw data with curves
vbgf_k_plot<-
cso_dat_lfm %>%
  group_by(treatment) %>%
  add_fitted_draws(vbgf_mod_k_only) %>% #re_formula = NA to deal with random effects
  ggplot(aes(x = age_final, y = fork_length_cm, color = treatment, fill = treatment,  alpha = 0.6)) +
  geom_point(data = cso_dat_lfm, size = 3, aes(shape = ordered(treatment)))+
  stat_lineribbon(aes(y = exp(.value)), .width=0) + #changed .width from 0 to have some error bars
   scale_fill_manual(values = c("#005AB5", "#DC3220")) + 
  scale_color_manual(values = c("#005AB5", "#DC3220")) +   
  scale_shape_manual(values = c(21,23)) + 
  theme_bw() + 
  xlab("Age (years)")+
  ylab("Length (cm)") +
  #  ggtitle("Only k varies")+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
vbgf_k_plot



##comparing raw data with curves from both models on same plot
dat_k_mod<-
  cso_dat_lfm %>%
  group_by(treatment) %>%
  add_fitted_draws(vbgf_mod_k_only)

dat_both_mod<-
  cso_dat_lfm %>%
  group_by(treatment) %>%
  add_fitted_draws(vbgf_mod_both)
  
vbgf_2_plot<-
  ggplot(data = dat_k_mod, aes(x = age_final, y = fork_length_cm, color = treatment, fill = treatment,  alpha = 0.6)) +
  geom_point(data = cso_dat_lfm, size = 3, aes(shape = ordered(treatment)))+
  stat_lineribbon(aes(y = exp(.value)), .width=.95) + #changed .width from 0 to have some error bars
    stat_lineribbon(data = dat_both_mod, aes(y = exp(.value)), .width=.95, lty=2) + #changed .width from 0 to have some error bars
   scale_fill_manual(values = c("#005AB5", "#DC3220")) + 
  scale_color_manual(values = c("#005AB5", "#DC3220")) +   
  scale_shape_manual(values = c(21,23)) + 
  theme_bw() + 
  xlab("Age (years)")+
  ylab("Length (cm)") +
  #  ggtitle("Only k varies")+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
vbgf_2_plot
##yup, nearly identical.

```


##Combine plots for paper
```{r}
vbgf_comb_plot <-
  ggdraw() + 
  draw_plot(vbgf_k_plot, x = 0, y = 0, width = 0.5, height = 1) +
  draw_plot(vbgf_k_post_plot,  x = 0.5, y = 0, width = 0.5, height = 1) 
vbgf_comb_plot

```


###respone to reviewers: run re-parametized VBGF model in bayesian framework to estimate length at specific ages (model built up slowly in previous files)
```{r}
#model (extracted formula from fishmethods package):
# L ~ lphi + (lpsi - lphi) * (1 - ((lpsi - lchi)/(lchi - lphi))^(2 *(x$age - agephi)/(agepsi - agephi)))/(1 - ((lpsi - lchi)/(lchi - lphi))^2)
###we're estimating lphi, lchi, and lpsi. agepsi and agephi will be set (phi = 2, psi = 6 - to cover age range with most data).


##split into rats/no rats for each age (lphi, lpsi, lchi)
repram_prior_2<- c(
  prior(uniform(15, 27), nlpar = "lphi", lb = 15, ub = 27), #constrained between approx minimum and maximum size
  prior(uniform(15, 27), nlpar = "lpsi", lb = 15, ub = 27), #constrained between approx minimum and maximum size
  prior(uniform(15, 27), nlpar = "lchi", lb = 15, ub = 27), #constrained between approx minimum and maximum size
  prior(normal(0,10), nlpar = "lphi1"), #allow lphi to vary by rat status in either direction
  prior(normal(0,10), nlpar = "lpsi1"), #allow lphi to vary by rat status in either direction
  prior(normal(0,10), nlpar = "lchi1") #allow lphi to vary by rat status in either direction
)


reparam_mod_treat2 <- 
  brm(
    bf(
      llength ~ log((lphi+lphi1*tin) + ((lpsi+lpsi1*tin) - (lphi+lphi1*tin)) * 
        (1 - (((lpsi+lpsi1*tin)-(lchi+lchi1*tin))/((lchi+lchi1*tin)-(lphi+lphi1*tin)))^(2 *(age_final - 2)/(6 - 2)))
      /(1 - (((lpsi+lpsi1*tin)-(lchi+lchi1*tin))/((lchi+lchi1*tin)-(lphi+lphi1*tin)))^2)),
     lphi~1, lpsi~1, lchi~1, lphi1~1, lpsi1~1, lchi1~1, 
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = repram_prior_2,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "reparam_mod_treat2"
  )

print(reparam_mod_treat2) 
plot(reparam_mod_treat2, ask = FALSE)

#do manual pp_check:
y <- cso_dat_lfm$llength
yrep <- posterior_predict(reparam_mod_treat2) %>%
  na.exclude()
ppc_dens_overlay(y, yrep[1:10,])
###all looks good


hypothesis(reparam_mod_treat2, c("lphi1_Intercept  > 0", ##0.90
                                  "lpsi1_Intercept  > 0", ##0.6
                                  "lchi1_Intercept  > 0")) #1.00
#strong evidence for differences in lphi and lchi


#Compile posterior distributions:
reparam_post <- posterior_samples(reparam_mod_treat2)
median_hdi(reparam_post$b_lphi1_Intercept, .width = .95) ## 0.5843224	-0.3702454	1.494105		
median_hdi(reparam_post$b_lpsi1_Intercept, .width = .95) ## 0.2290143	-1.489648	2.140322
median_hdi(reparam_post$b_lchi1_Intercept, .width = .95) ## 2.240697	0.9637946	3.472135	

median_hdi(reparam_post$b_lphi_Intercept, .width = .95) ## 19.40771	18.79999	20.02697			
median_hdi(reparam_post$b_lpsi_Intercept, .width = .95) ## 23.14062	21.92997	24.47871	
median_hdi(reparam_post$b_lchi_Intercept, .width = .95) ## 21.30782	20.58563	21.9651	



est_age_length<-
  reparam_mod_treat2 %>%
  spread_draws(b_lphi_Intercept, b_lpsi_Intercept, b_lchi_Intercept, b_lphi1_Intercept, b_lpsi1_Intercept, b_lchi1_Intercept) %>%
   median_qi(lphi_rat = b_lphi_Intercept,
            lphi_bird = b_lphi_Intercept + b_lphi1_Intercept,
             lchi_rat = b_lchi_Intercept, 
            lchi_bird = b_lchi_Intercept + b_lchi1_Intercept,
             lpsi_rat = b_lpsi_Intercept, 	
             lpsi_bird = b_lpsi_Intercept + b_lpsi1_Intercept) 
est_age_length


est_age_length_75<-
  reparam_mod_treat2 %>%
  spread_draws(b_lphi_Intercept, b_lpsi_Intercept, b_lchi_Intercept, b_lphi1_Intercept, b_lpsi1_Intercept, b_lchi1_Intercept) %>%
   median_qi(lphi_rat = b_lphi_Intercept,
            lphi_bird = b_lphi_Intercept + b_lphi1_Intercept,
             lchi_rat = b_lchi_Intercept,
            lchi_bird = b_lchi_Intercept + b_lchi1_Intercept,
             lpsi_rat = b_lpsi_Intercept,
             lpsi_bird = b_lpsi_Intercept + b_lpsi1_Intercept, .width=.75)
est_age_length_75


##use above lengths at age to calculate grazing rates based on Lange et al. 2020 rates:
#	             a	       b
#Grazing: 	  0.0433	2.209
#Bioerosion:	0.0131	2.211
#rate = a*TL^b 	


```


###Plots for reparameterized vbgf - supplement
```{r}

##plot as one:
reparam_post_plot_all <-
reparam_mod_treat2 %>%
  spread_draws(b_lchi1_Intercept, b_lpsi1_Intercept, b_lphi1_Intercept) %>%
  ggplot(aes(fill = stat(x < 0))) +
  stat_halfeyeh(aes(y = 0, x = b_lpsi1_Intercept, ), point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) +
    stat_halfeyeh(aes(y = 1, x = b_lchi1_Intercept, ), point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) +
    stat_halfeyeh(aes(y = 2, x = b_lphi1_Intercept, ), point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) +
  scale_fill_manual(values = alpha(c("#005AB5", "#DC3220"),0.3))+
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() + 
  xlab("Difference in length-at-age (cm)")+ 
  ylab("") +
  scale_y_continuous(breaks=c(0,1,2), labels= c("0" = "age 6", "1" = "age 4", "2" = "age 2")) +
  scale_x_continuous(limits=c(-4, 5), breaks=seq(-4, 5, 1))+
  #scale_x_continuous(limits=c(-0.05, 0.25), breaks=seq(0, 0.25,.1))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
reparam_post_plot_all


#ggsave(filename = "reparam_post_plot_all_length_age_diff_16march.pdf", 
#       plot = reparam_post_plot_all,
#       width = 15, 
#       height = 10,
#       units = "cm",
#       dpi = 300,
#       family = "sans",
#       useDingbats=FALSE)



```


###response to reviewers - run models for difference in max age and max length between rat-free and rat-infested islands: 
```{r}
#calculated #mean of upper quartile calculations to get max age and max length by island - saved in:
load("cso_max_length_age_data.Rdata")

str(max_length_age_dat)

#run model - length
range(max_length_age_dat$mean_length) #21.06667 26.30000
mean(max_length_age_dat$mean_length) #23.49063

max_length_mod <- 
  brm(data = max_length_age_dat, family = gaussian,
      mean_length  ~ treatment_b + (1|atoll),
      prior = c(prior(normal(23,5), class = Intercept), ###weakly informative prior based on mean and range
                prior(normal(0, 10), class = "b", coef = "treatment_bbirdy")), #weakly informative prior
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.999, max_treedepth = 15), 
      sample_prior="yes",
      file = "max_length_mod")

print(max_length_mod) #all good
plot(max_length_mod, ask = FALSE)
pp_check(max_length_mod, nsamples=100) # a bit messy, but not bad

hypothesis(max_length_mod, "treatment_bbirdy>0") # 98% posterior prob


#run model - age
range(max_length_age_dat$mean_age)
mean(max_length_age_dat$mean_age)

max_age_mod <- 
  brm(data = max_length_age_dat, family = gaussian,
      mean_age  ~ treatment_b + (1|atoll),
      prior = c(prior(normal(6,2), class = Intercept), ###weakly informative prior based on mean and range
                prior(normal(0, 10), class = "b", coef = "treatment_bbirdy")), #weakly informative prior
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.999, max_treedepth = 15), 
      sample_prior="yes",
      file = "max_age_mod")

print(max_age_mod)
plot(max_age_mod, ask = FALSE)
pp_check(max_age_mod, nsamples=100) #not great, but not terrible. 

hypothesis(max_age_mod, "treatment_bbirdy>0") #0.95



#Compile posterior distributions:
age_post <- posterior_samples(max_age_mod)
median_hdi(age_post$b_treatment_bbirdy, .width = .95) ## 1.358834	-0.3977543	3.068403


length_post <- posterior_samples(max_length_mod)
median_hdi(length_post$b_treatment_bbirdy, .width = .95) ## 1.933577	-0.0008097673	3.736939


est_max_age<-
  max_age_mod %>%
  spread_draws(b_treatment_bbirdy, b_Intercept) %>%
   median_qi(max_age_rat = b_Intercept, #5.298061	
            max_age_bird = b_treatment_bbirdy + b_Intercept) #6.644844
est_max_age


est_max_length<-
  max_length_mod %>%
  spread_draws(b_treatment_bbirdy, b_Intercept) %>%
   median_qi(max_length_rat = b_Intercept, #22.09622
            max_length_bird = b_treatment_bbirdy + b_Intercept) #24.01752
est_max_length


##percent difference
max_age_mod%>%
  spread_draws(b_Intercept, b_treatment_bbirdy) %>%
   median_qi(rat = b_Intercept, 
            bird_diff = b_treatment_bbirdy)%>%
   mutate(percent_diff = (bird_diff)/((rat+rat+bird_diff)/2)*100, #22.73256
          percent_change_rat_to_bird = (bird_diff)/rat*100) #25.64775


max_length_mod%>%
  spread_draws(b_Intercept, b_treatment_bbirdy) %>%
   median_qi(rat = b_Intercept, 
            bird_diff = b_treatment_bbirdy)%>%
   mutate(percent_diff = (bird_diff)/((rat+rat+bird_diff)/2)*100, #8.38389
          percent_change_rat_to_bird = (bird_diff)/rat*100) #8.750715

```


##posterior plots from max age/length models:
```{r}

post_plot_maxage <-
max_age_mod %>%
  spread_draws(b_treatment_bbirdy) %>%
  ggplot(aes(y = 0, x = b_treatment_bbirdy, fill = stat(x < 0))) +
  stat_halfeyeh(point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) + #.width=c(.95, 0.75)
  scale_fill_manual(values = alpha(c("#005AB5", "#DC3220"),0.3))+
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() + 
  xlab("Difference in maximum age")+ 
    ylab("Posterior density") +
    scale_x_continuous(limits = c(-3.8,8.6), breaks=seq(-2, 6, 2))+
#  ylab("") +
  #ggtitle("Difference in length at age 4")+
  #scale_x_continuous(limits=c(-4, 5), breaks=seq(-4, 5, 1))+
  #scale_x_continuous(limits=c(-0.05, 0.25), breaks=seq(0, 0.25,.1))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
post_plot_maxage



post_plot_maxlength <-
max_length_mod %>%
  spread_draws(b_treatment_bbirdy) %>%
  ggplot(aes(y = 0, x = b_treatment_bbirdy, fill = stat(x < 0))) +
  stat_halfeyeh(point_interval=median_hdi, .width=c(.95, .75), fatten_point = 2, slab_alpha = 0.6) + # .width=c(.95, 0.75)
  scale_fill_manual(values = alpha(c("#005AB5", "#DC3220"),0.3))+
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() + 
  xlab("Difference in maximum length")+ 
  ylab("Posterior density") +
  #ggtitle("Difference in length at age 4")+
    scale_x_continuous(limits = c(-3.8,8.6), breaks=seq(-2, 6, 2))+
  #scale_x_continuous(limits=c(-0.05, 0.25), breaks=seq(0, 0.25,.1))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16),
         axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
post_plot_maxlength

```

###box plots for max age/length
```{r}

max_age_box<-
max_length_age_dat %>% 
 mutate(treatment = factor(treatment, levels=c("ratty", "birdy")))  %>% #re-arrange to match posterior plots
  ggplot(aes(x=treatment, y=mean_age, fill=treatment, shape = treatment, alpha = 0.6)) + 
  geom_boxplot(outlier.size=0, outlier.stroke=0) +
  geom_point(stat="identity", size = 2, position = position_jitterdodge(dodge.width=.7)) + 
  scale_fill_manual(values = c("#DC3220", "#005AB5"))+
  scale_shape_manual(values = c(23,21)) + 
  scale_x_discrete(labels=c("rat-infested", "rat-free")) +
 # scale_y_continuous(labels=scales::number_format(accuracy=0.1))+
  theme_bw() + 
  xlab(" ")+
  ylab("Maximum age (years)") +
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
max_age_box


max_length_box<-
max_length_age_dat %>% 
 mutate(treatment = factor(treatment, levels=c("ratty", "birdy")))  %>% #re-arrange to match posterior plots
  ggplot(aes(x=treatment, y=mean_length, fill=treatment, shape = treatment, alpha = 0.6)) + 
  geom_boxplot(outlier.size=0, outlier.stroke=0) +
  geom_point(stat="identity", size = 2, position = position_jitterdodge(dodge.width=.7)) + 
  scale_fill_manual(values = c("#DC3220", "#005AB5"))+
  scale_shape_manual(values = c(23,21)) + 
  scale_x_discrete(labels=c("rat-infested", "rat-free")) +
 # scale_y_continuous(labels=scales::number_format(accuracy=0.1))+
  theme_bw() + 
  xlab(" ")+
  ylab("Maximum length (cm)") +
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
max_length_box


```

###combine max age/length plots and save:
```{r}
####NEW:
max_age_plot <-
  ggdraw() + 
  draw_plot(max_age_box, x = 0, y = 0, width = 0.5, height = 1) +
  draw_plot(post_plot_maxage,  x = 0.5, y = 0, width = 0.5, height = 1) 
max_age_plot

max_length_plot <-
  ggdraw() + 
  draw_plot(max_length_box, x = 0, y = 0, width = 0.5, height = 1) +
  draw_plot(post_plot_maxlength,  x = 0.5, y = 0, width = 0.5, height = 1) 
max_length_plot

```



##Combine growth plots with GSI plots from other file: 
```{r}

###new - to make it aligned:
vbgf_k_plot
vbgf_k_post_plot
max_length_box
post_plot_maxlength
max_age_box
post_plot_maxage
gsi_boxplot
gsi_post_plot


aligned_raw_plots_1<-plot_grid(vbgf_k_plot, max_age_box, max_length_box, gsi_boxplot, ncol = 1, align = "hv")
aligned_raw_plots_1

aligned_post_1<-plot_grid(vbgf_k_post_plot, post_plot_maxage, post_plot_maxlength, gsi_post_plot, ncol = 1, align = "hv")
aligned_post_1


comb_plot_1 <-
  ggdraw() + 
  draw_plot(aligned_raw_plots_1, x = 0, y = 0, width = 0.5, height = 1) + # draw_plot(aligned_raw_plots, x = 0, y = 0, width = 0.55, height = 1)
  draw_plot(aligned_post_1,  x = 0.5, y = 0, width = 0.5, height = 1) #  draw_plot(post_plots,  x = 0.55, y = 0, width = 0.45, height = 1) 
comb_plot_1


ggsave(filename = "fig1_comb_March16.pdf", 
       plot = comb_plot_1,
       width = 20, 
       height = 25,
       units = "cm",
       dpi = 300,
       family = "sans",
       useDingbats=FALSE)

```





###run overall vbgf and allow to vary by individual to extract data - corr between growth and gsi
```{r}
#see Zhang et al. 2009 and Scherrer et al. 2021

##run model - neither----------------
vbgf_prior_none<- c(
  prior(uniform(15.2, 53.6), nlpar = "Linf", lb = 15.2, ub = 53.6), #constraind between minimum size and twice maximum
  prior(uniform(0,2), nlpar = "K", lb=0, ub = 2), #constrained because must be > 0, and k not > 2
  prior(uniform(0, 15.2), nlpar = "L0", lb = 0, ub = 15.2) ###L0 constrained bewettn 0 and minimum size observed
)


vbgf_mod_none_ind <- 
  brm(
    bf(
      llength ~ log(Linf-(Linf-L0)*exp(-(K)*age_final)),
      Linf~1, K~1 + (1|sample_ID) , L0~1,
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = vbgf_prior_none,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "vbgf_mod_none_ind"
  )
print(vbgf_mod_none_ind) 


plot(vbgf_mod_none_ind, ask = FALSE)
pp_check(vbgf_mod_none_ind)
###alll look sokay

###ok extract individual estimates: 
ind_K<-
  vbgf_mod_none_ind %>%
  spread_draws(b_K_Intercept, r_sample_ID__K[sample_ID,]) %>%
   median_qi(k_median = b_K_Intercept + r_sample_ID__K,
             k_diff = r_sample_ID__K)
ind_K

###nice, okay now regress these against predictor variables: 
cso_ind_K_dat<-inner_join(ind_K, cso_dat_lfm, by=c("sample_ID")) 
cso_ind_K_dat

ind_K_dat$sample_ID <- as_factor(ind_K_dat$sample_ID)


##run model with just treatment to start: 
ind_k_atoll <- 
  brm(data = cso_ind_K_dat, family = gaussian,
      k_median  ~ treatment_b  + (1|atoll),
       prior = c(prior(normal(0.37,10), class = Intercept), #weakly informative intercept prior
                prior(normal(0, 10), class = "b", coef = "treatment_bbirdy")), #weakly informative rat prior
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.999, max_treedepth = 15), 
      sample_prior="yes",
      file = "ind_k_atoll")

print(ind_k_atoll) #16 divergent trans
plot(ind_k_atoll, ask=FALSE)
pp_check(ind_k_atoll)


hypothesis(ind_k_atoll, "treatment_bbirdy>0") ##100% posterior prob
##still shows birdy effect



####write data to file:
cso_ind_K_dat
####save as Rdata:
#save(cso_ind_K_dat, file = "cso_ind_K_dat.Rdata")
load("cso_ind_K_dat.Rdata")

```


###look for corr between k_diff, GSI at individual level:
```{r}
library(correlation) #https://rdrr.io/cran/correlation/man/cor_test.html
#install.packages('BayesFactor')
## when using cor_test() if you set multilevel=TRUE in it, the resulting correlations are partial one.


##create reduced data frame with only variables we're interested in:
cso_ind_K_dat_red<-
  cso_ind_K_dat%>% 
  select(c("k_median", "GSI", "atoll"))

cor_test(cso_ind_K_dat_red, "k_median", "GSI", multilevel = TRUE) #, bayesian = TRUE, partial_bayesian = TRUE
#Parameter1 | Parameter2 |     r |         95% CI | t(92) |       p
#------------------------------------------------------------------
#k_median   |        GSI | -0.28 | [-0.46, -0.08] | -2.80 | 0.006**

cor_test(cso_ind_K_dat_red, "k_median", "GSI", multilevel = TRUE, bayesian = TRUE) #, partial_bayesian = TRUE
#Parameter1 | Parameter2 |   rho |         95% CI |       pd | % in ROPE |    BF |         Prior
#-----------------------------------------------------------------------------------------------
#k_median   |        GSI | -0.26 | [-0.41, -0.11] | 99.60%** |     4.90% | 8.07* | Beta (3 +- 3)



##remove k outlier and make sure pattern holds:
cso_ind_K_dat_red2<-
  cso_ind_K_dat%>% 
  select(c("k_median", "GSI", "atoll")) %>% 
  filter(k_median>=0.2)

cor_test(cso_ind_K_dat_red2, "k_median", "GSI", multilevel = TRUE) #trend holds
#Parameter1 | Parameter2 |     r |         95% CI | t(91) |      p
#-----------------------------------------------------------------
#k_median   |        GSI | -0.22 | [-0.41, -0.02] | -2.17 | 0.033*

cor_test(cso_ind_K_dat_red2, "k_median", "GSI", multilevel = TRUE, bayesian = TRUE) # still quite strong
#Parameter1 | Parameter2 |   rho |         95% CI |      pd | % in ROPE |   BF |         Prior
#---------------------------------------------------------------------------------------------
#k_median   |        GSI | -0.21 | [-0.37, -0.06] | 97.70%* |    12.47% | 2.06 | Beta (3 +- 3)

```




##old: VBGF model with k varying by atoll
```{r}
#need to specify priors for nl models
#based on priors in Graham et al. 2018 (who used same formula)
vbgf_prior<- c(
  prior(uniform(15.2, 53.6), nlpar = "Linf", lb = 15.2, ub = 53.6), #constraind between minimum size and twice maximum
  prior(uniform(0,2), nlpar = "K", lb=0, ub = 2), #constrained because must be > 0, and k not > 2
  prior(uniform(0, 15.2), nlpar = "L0", lb = 0, ub = 15.2), ###L0 constrained bewettn 0 and minimum size observed
  prior(normal(0, 1), nlpar = "K1") ###weak, to allow either positive or negative (or 0) birdy effect
)

#run model
vbgf_mod <- 
  brm(
    bf(
      llength ~ log(Linf-(Linf-L0)*exp(-(K+K1*tin)*age_final)),
      Linf~1, K~1+(1|atoll), L0~1, K1~1,
      nl=TRUE),
    data = cso_dat_lfm, family = gaussian,
    prior = vbgf_prior,
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
    control = list(adapt_delta = 0.9999, max_treedepth=15), 
    sample_prior=TRUE,
      file = "vbgf_mod"
  )
print(vbgf_mod) 
plot(vbgf_mod, ask = FALSE)
pp_check(vbgf_mod)
pairs(vbgf_mod)
#14 divergent transitions, but ess, and other diagnostics look good
prior_summary(vbgf_mod)

#compare posterior vs. prior distributions
#Linf:
posterior_samples(vbgf_mod) %>% 
  select(b_Linf_Intercept, prior_b_Linf) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#K:
posterior_samples(vbgf_mod) %>% 
  select(b_K_Intercept, prior_b_K) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope (rat):
posterior_samples(vbgf_mod) %>% 
  select(b_L0_Intercept, prior_b_L0) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope (rat):
posterior_samples(vbgf_mod) %>% 
  select(b_K1_Intercept, prior_b_K1) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()


#Compile posterior distributions:
vbgf_post <- posterior_samples(vbgf_mod)
median_hdi(vbgf_post$b_K1_Intercept, .width = .95) #median = 0.08, min = 0.021, max = 0.14


#test hypothesis that birdy effect is > 0
hypothesis(vbgf_mod, "K1_Intercept  > 0") ##evidence ratio = 469.59, posterior prob = 1


est_k<-
  vbgf_mod %>%
  spread_draws(b_K1_Intercept, b_K_Intercept) %>%
   median_qi(k_rat = b_K_Intercept, #0.2769853	0.05928513	0.9180025
            k_bird = b_K_Intercept + b_K1_Intercept) #0.3565911	0.1341369	0.9994793
est_k

est_k2<-
  vbgf_mod_k_only %>%
  spread_draws(b_K1_Intercept, b_K_Intercept) %>%
   median_qi(k_rat = b_K_Intercept, #0.2747117	0.1709674	0.4387155	
            k_bird = b_K_Intercept + b_K1_Intercept) #0.3752644	0.2231286	0.6074888	
est_k2

##very similar results as model without atoll, so use simpler model (no atoll) to match Graham et al. 2018. 

```

