---
title: "Estimated population-level fecundity"
output: html_document
---

#Step 0: Load packages
```{r}
library(tidyverse)
library(brms)
library(tidybayes)
library(modelr) #for data_grid function
library(bayesplot)
library(cowplot) #for plots

```


#Load gonad weight-length data
```{r}

load("cso_data.Rdata")

##want to use lagoon, female, stage 3:
str(cso_dat_lfm)

##re-level to make ratty the reference level
cso_dat_lfm$treatment_b<-relevel(cso_dat_lfm$treatment, "ratty")

```


#Re-load model results for relationships between gonad weight, length, and bird vs. rat islands
```{r}
gw_length_mod <- 
  brm(data = cso_dat_lfm, family = gaussian,
      gonad.weight_g  ~ c_length+treatment_b + (1|atoll),
      prior = c(prior(normal(2,10), class = Intercept), ###weakly informative prior
                prior(lognormal(0, 1), class = "b", coef = "c_length"), #this has to be above 0 (positive relationship between gonad weight-length)
                prior(normal(0, 10), class = "b", coef = "treatment_bbirdy")), #weakly informative prior
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.999, max_treedepth = 15), 
      sample_prior="yes",
      file = "gw_length_mod")
print(gw_length_mod)

##RE-RUN WITHOUT INTERCEPT TO MAKE EASIER TO EXTRACT EFFECTS:
gw_length_mod_no_int <- 
  brm(data = cso_dat_lfm, family = gaussian,
      gonad.weight_g  ~ 0 + treatment_b+ c_length + (1|atoll),
      prior = c(
                prior(lognormal(0, 1), class = "b", coef = "c_length")), #this has to be above 0 (positive relationship between gonad weight-length)
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.999, max_treedepth = 15), 
      sample_prior="yes",
      file = "gw_length_mod_no_int")
print(gw_length_mod_no_int)
##estimates match above model

```


#Load size distribution data and combine with estimated gonad weights
```{r}
load("cso_surveys_data.Rdata")

##want sordidus only, 2018 - by individual so can get lengths
str(cso_only_2018)

##center length and change variable names to match cso_dat_lfm for getting gonad weights from model
cso_only_2018_recode <- 
  cso_only_2018 %>%
  mutate(c_length = Length - mean(Length),
         atoll= recode(Atoll, Great_Chagos_Bank="Great Chagos Bank", Peros_Banhos = "Peros Banhos"),
         treatment_b = relevel(Treatment, "Rats"),
         treatment_b=recode(treatment_b, Rats = "ratty", No_rats = "birdy"))


cso_only_2018_mod <-
cso_only_2018_recode %>%
  group_by(atoll, treatment_b) %>%
  data_grid(c_length = c_length) %>%  ##rather than using seq_by, just use exact lengths in survey data to get estimated gonad weights
  add_fitted_draws(gw_length_mod_no_int) %>%
  median_hdi()

#combine data
cso_only_2018_re_mod<-inner_join(cso_only_2018_mod, cso_only_2018_recode, by = c("atoll", "treatment_b", "c_length"), all.x=TRUE)
cso_only_2018_re_mod

```



#Limit data to females only
```{r}
#Size-at-sex-change from published studies:

#Barba et al. (master's thesis): females reach maturity (50%) between 113.4 and 157.8 mm FL
#males reach maturity between 173.3 and 223.8 mm FL

#NOAA Marianas Island report: females reach 50% maturity at 148 mm TL, length at 50% sex change for males = 207 mm TL
#BUT some females never undergo sex change, so could be large females on the surveys....


#DeMartini and Howard 2016 (Chlorurus spilurus, formely sordidus - sister species in Pacific) in Hawaii:
#Length 50% maturity: 17.2, length 50% sex change: 27.3 cm (fork lengths)
##DeMartini and Howard also have good summary table of size at maturity and size at sex change!  includes Gust and Noaa report. 

##Gust 2004: differences in size and age at sex change just 20 km apart (outer shelf reef crest versus mid-shelf habitats)

##From 2019 BIOT collections, all individuals > 26.9 cm were male...so this seems like a reasonable cut-off
##Smallest collected was 14.0 cm - all were mature

##Could run one model with smallest sizes (e.g., 10 cm maturity size) and one with largest sizes (e.g., 18 cm length maturity) so can bound results

##so using these cut-offs, assign individuals to female or male

##make new df with only females: 
cso_only_2018_sex<-
  cso_only_2018_re_mod  %>%
  mutate(sex = case_when(
    Length <10 ~ "Immature",
    Length >=10 & Length<27 ~ "Female",
    Length >=27 ~ "Male"
  ))

cso_only_2018_F<-
   cso_only_2018_sex  %>%
  filter(sex=="Female")

cso_only_2018_F
##if cut-off = 14, then only 32 females - still no effect, but more trend towards positive birdy effect
##if cut-off = 10, then 119 rows - no effect, but more trend towards positive ratty effect. 

```


###combine survey data and gonad weight data, sum gonad weight by transect
```{r}
str(cso_only_2018_F)

##sum by transect 
cso_only_sum_2018_gw<-
  cso_only_2018_F  %>%
  group_by(Island, Transect, .drop=FALSE)  %>%  
  summarize(sum_bio=sum(Biomass),
            sum_dens = sum(Abundance),
            sum_gonad=sum(.value),
            sum_lower=sum(.lower),
            sum_upper=sum(.upper))%>%
filter(!(Island == "Nelson_Island")) #drop Nelson 2018 because not surveyed
str(cso_only_sum_2018_gw)
levels(cso_only_sum_2018_gw$Island) ##missing South Brother, because no female sordidus there. 

###add back in treatment data:
cso_only_sum_2018_gw2<-inner_join(cso_only_sum_2018_gw, cso_only_sum[2:4], by = c("Island"), all.x=TRUE)
cso_only_sum_2018_gw2<-unique(cso_only_sum_2018_gw2)
cso_only_sum_2018_gw2$Island<-as.factor(cso_only_sum_2018_gw2$Island)
nrow(cso_only_sum_2018_gw2)
levels(cso_only_sum_2018_gw2$Island)


###add in 0 rows for South Brother
new <- as.data.frame(cbind(
Island = c("South_Brother","South_Brother","South_Brother","South_Brother"),
Transect=c("1", "2", "3", "4"),
Treatment=c("No_rats", "No_rats","No_rats","No_rats"),
Atoll=c("Great_Chagos_Bank", "Great_Chagos_Bank", "Great_Chagos_Bank", "Great_Chagos_Bank" ),
sum_bio = c(0,0,0,0),
sum_dens = c(0,0,0,0),
sum_gonad = c(0,0,0,0),
sum_lower = c(0,0,0,0),
sum_upper = c(0,0,0,0)),  stringsAsFactors=FALSE)
new$sum_bio<-as.numeric(new$sum_bio)
new$sum_dens<-as.numeric(new$sum_dens)
new$sum_gonad<-as.numeric(new$sum_gonad)
new$sum_lower<-as.numeric(new$sum_lower)
str(new)

cso_only_sum_2018_gw3<-bind_rows(data.frame(cso_only_sum_2018_gw2), new)
str(cso_only_sum_2018_gw3)
cso_only_sum_2018_gw3$Island<-as.factor(cso_only_sum_2018_gw3$Island)
cso_only_sum_2018_gw3$Treatment<-as.factor(cso_only_sum_2018_gw3$Treatment)
cso_only_sum_2018_gw3$Atoll<-as.factor(cso_only_sum_2018_gw3$Atoll)
str(cso_only_sum_2018_gw3)

```


##run model
```{r}
###run model: 
###re-level for birdy effect: 
cso_only_sum_2018_gw3$treatment_b<-relevel(cso_only_sum_2018_gw3$Treatment, "Rats")
mean(cso_only_sum_2018_gw3$sum_gonad)

###try hurdle gamma:
popn_gonad_mod_hg_2 <- 
  brm(data = cso_only_sum_2018_gw3, family  = hurdle_gamma(link="log"),
      sum_gonad  ~ treatment_b  + (1|Atoll),
      prior = c(prior(normal(0,1), class = "b"), ##weakly informative, allow to vary, sd = 1 bc log link
                prior(normal(2,1), class = "Intercept")), ##weakly informative, mean = 2 bc log(mean(gw))
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.9999), 
      sample_prior="yes",
      file = "popn_gonad_mod_hg_2")
print(popn_gonad_mod_hg_2) 
plot(popn_gonad_mod_hg_2) 
pp_check(popn_gonad_mod_hg_2) #looks fine


#intercept:
posterior_samples(popn_gonad_mod_hg_2) %>% 
  select(b_Intercept, prior_Intercept) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope:
posterior_samples(popn_gonad_mod_hg_2) %>% 
  select(b_treatment_bNo_rats, prior_b) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#Compile posterior distributions:
popn_gonad_post <- posterior_samples(popn_gonad_mod_hg_2)
median_hdi(popn_gonad_post$b_treatment_bNo_rats, .width = .95) #median estimate = -0.37, ymin = -1.11, ymax = 0.49


#test hypothesis that birdy effect is > 0
hypothesis(popn_gonad_mod_hg_2, "treatment_bNo_rats>0") 
#evidence ratio = 0.22, posterior probability = 0.18



###re-run with lower and upper estimates:
popn_gonad_mod_hg_2_lower <- 
  brm(data = cso_only_sum_2018_gw3, family  = hurdle_gamma(link="log"),
      sum_lower  ~ treatment_b  + (1|Atoll),
       prior = c(prior(normal(0,1), class = "b"), ##weakly informative, allow to vary, sd = 1 bc log link
                prior(normal(2,1), class = "Intercept")), ##weakly informative, mean = 2 bc log(mean(gw))
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.9999), 
      sample_prior="yes",
      file = "popn_gonad_mod_hg_2_lower")
print(popn_gonad_mod_hg_2_lower)
pp_check(popn_gonad_mod_hg_2_lower)
hypothesis(popn_gonad_mod_hg_2_lower, "treatment_bNo_rats>0") 
##evidence ratio = 0.14     posterior prob = 0.12  


popn_gonad_mod_hg_2_upper <- 
  brm(data = cso_only_sum_2018_gw3, family  = hurdle_gamma(link="log"),
      sum_lower  ~ treatment_b  + (1|Atoll),
      prior = c(prior(normal(0,1), class = "b"), ##weakly informative, allow to vary, sd = 1 bc log link
                prior(normal(2,1), class = "Intercept")), ##weakly informative, mean = 2 bc log(mean(gw))
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.9999), 
      sample_prior="yes",
      file = "popn_gonad_mod_hg_2_upper")
print(popn_gonad_mod_hg_2_upper)
hypothesis(popn_gonad_mod_hg_2_upper, "treatment_bNo_rats>0") 
##similar to lower estimate: ER = 0.13    Posterior prob =   0.12  

###so all of these results are similar regardless of using point estimate, or high or low 95% HPDI for gonad weight

```


##Plot: boxplots of raw data
```{r}

popn_gonad_boxplot_no0<-
  cso_only_sum_2018_gw3 %>% 
  filter(sum_gonad>0)%>% ##exclude zeros because hurdle model
  mutate(Treatment = factor(Treatment, levels=c("Rats", "No_rats")))  %>% #re-arrange to match posterior plots
  ggplot(aes(x=Treatment, y=sum_gonad, fill=Treatment, alpha = 0.6, shape = Treatment)) + 
  geom_boxplot(outlier.size=0, outlier.stroke=0) +
  geom_point(stat="identity", size = 2, position = position_jitterdodge(dodge.width=.1)) + 
  scale_fill_manual(values = c("#DC3220", "#005AB5"))+
  scale_shape_manual(values = c(23,21)) + 
  scale_x_discrete(labels=c("rat-infested", "rat-free")) +
  scale_y_continuous(trans='log10')+ #plot on log scale because log link in model #, limits=c(0, 1000)
  theme_bw() + 
  xlab(" ")+
  ylab("Gonad weight (g/ha)") +
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.text.x = element_text(size=16))
popn_gonad_boxplot_no0


nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="Rats"&cso_only_sum_2018_gw3$sum_gonad>0,]) ##13
nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="Rats",]) ##24
nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="Rats"&cso_only_sum_2018_gw3$sum_gonad==0,]) ##11

nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="No_rats"&cso_only_sum_2018_gw3$sum_gonad>0,]) ##7
nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="No_rats",]) ##20
nrow(cso_only_sum_2018_gw3[cso_only_sum_2018_gw3$Treatment=="No_rats"&cso_only_sum_2018_gw3$sum_gonad==0,]) ##13

  

```



##Posterior plot
```{r}

popn_gonad_post_plot<-
popn_gonad_mod_hg_2 %>%
  spread_draws(b_treatment_bNo_rats) %>%
 # mutate(exp_b=exp(b_treatment_bNo_rats))%>%
  ggplot(aes(y = 0, x = b_treatment_bNo_rats, fill = stat(x < 0))) +
  stat_halfeyeh(point_interval=median_hdi, .width=.95, fatten_point = 2, slab_alpha = 0.6) +
  scale_fill_manual(values = alpha(c("#005AB5", "#DC3220"),0.3))+
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() + 
  xlab("Gonad weight")+
  ylab("") +
  scale_y_continuous(NULL, breaks = NULL) +
 # scale_x_continuous(limits=c(-2.5, 2.5), breaks=seq(-2, 2,1))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), 
        legend.position = "none",  
        rect = element_rect(fill = "transparent"),  
        plot.background = element_rect(fill = "transparent", color = NA),
        text=element_text(size=16,  family="sans"),
        axis.title.y = element_text(size=16, face="bold"))
popn_gonad_post_plot

```


###combine plots and save
```{r}
#plots from other file for individuals:
ind_rep_plot <-
  ggdraw() + 
  draw_plot(gw_length_plot, x = 0, y = 0, width = 0.52, height = 1) +
  draw_plot(gwl_post_plot,  x = 0.52, y = 0, width = 0.47, height = 1) 
ind_rep_plot

#plots from this file:
popn_rep_plot <-
  ggdraw() + 
  draw_plot(popn_gonad_boxplot_no0, x = 0, y = 0, width = 0.52, height = 1) +
  draw_plot(popn_gonad_post_plot,  x = 0.52, y = 0, width = 0.47, height = 1) 
popn_rep_plot


gw_popn_ind_plots_comb<-plot_grid(ind_rep_plot, popn_rep_plot, ncol = 1, align = "v")
gw_popn_ind_plots_comb

ggsave(filename = "gw_popn_ind_plots_comb_aug6.pdf", 
       plot = gw_popn_ind_plots_comb,
       width = 20, 
       height = 15,
       units = "cm",
       dpi = 300,
       family = "sans",
       useDingbats=FALSE)
```




##EXTRAS: some more model comparisons to be sure results are robust:
##re-run using a different length cut-off - set-up:
```{r}
cso_only_2018_sex2<-
  cso_only_2018_re_mod  %>%
  mutate(sex = case_when(
    Length <14 ~ "Immature",
    Length >=14 & Length<27 ~ "Female",
    Length >=27 ~ "Male"
  ))

cso_only_2018_F2<-
   cso_only_2018_sex2  %>%
  filter(sex=="Female")

cso_only_2018_F2
##if cut-off = 14, then only 32 females - still no effect, but more trend towards positive birdy effect
##if cut-off = 10, then 119 rows - no effect, but more trend towards positive ratty effect. 


##sum by transect 
cso_only_sum_2018_gwb<-
  cso_only_2018_F2  %>%
  group_by(Island, Transect, .drop=FALSE)  %>%  
  summarize(sum_bio=sum(Biomass),
            sum_dens = sum(Abundance),
            sum_gonad=sum(.value))%>%
filter(!(Island == "Nelson_Island")) #drop Nelson 2018 because not surveyed
str(cso_only_sum_2018_gwb)
levels(cso_only_sum_2018_gwb$Island) ##missing South Brother, because no female sordidus there. 

###add back in treatment data:
cso_only_sum_2018_gw2b<-inner_join(cso_only_sum_2018_gwb, cso_only_sum[2:4], by = c("Island"), all.x=TRUE)
cso_only_sum_2018_gw2b<-unique(cso_only_sum_2018_gw2b)
cso_only_sum_2018_gw2b$Island<-as.factor(cso_only_sum_2018_gw2b$Island)
nrow(cso_only_sum_2018_gw2b)
levels(cso_only_sum_2018_gw2b$Island)


###add in 0 rows for South Brother
new <- as.data.frame(cbind(
Island = c("South_Brother","South_Brother","South_Brother","South_Brother"),
Transect=c("1", "2", "3", "4"),
Treatment=c("No_rats", "No_rats","No_rats","No_rats"),
Atoll=c("Great_Chagos_Bank", "Great_Chagos_Bank", "Great_Chagos_Bank", "Great_Chagos_Bank" ),
sum_bio = c(0,0,0,0),
sum_dens = c(0,0,0,0),
sum_gonad = c(0,0,0,0)),  stringsAsFactors=FALSE)
new$sum_bio<-as.numeric(new$sum_bio)
new$sum_dens<-as.numeric(new$sum_dens)
new$sum_gonad<-as.numeric(new$sum_gonad)
new$sum_lower<-as.numeric(new$sum_lower)
str(new)

cso_only_sum_2018_gw3b<-bind_rows(data.frame(cso_only_sum_2018_gw2b), new)
str(cso_only_sum_2018_gw3b)
cso_only_sum_2018_gw3b$Island<-as.factor(cso_only_sum_2018_gw3b$Island)
cso_only_sum_2018_gw3b$Treatment<-as.factor(cso_only_sum_2018_gw3b$Treatment)
cso_only_sum_2018_gw3b$Atoll<-as.factor(cso_only_sum_2018_gw3b$Atoll)
str(cso_only_sum_2018_gw3b)

```


##run model - different size cutoff
```{r}
###run model: 
###re-level for birdy effect: 
cso_only_sum_2018_gw3b$treatment_b<-relevel(cso_only_sum_2018_gw3b$Treatment, "Rats")
mean(cso_only_sum_2018_gw3b$sum_gonad)

###try hurdle gamma:
popn_gonad_mod_hg_2b <- 
  brm(data = cso_only_sum_2018_gw3b, family  = hurdle_gamma(link="log"),
      sum_gonad  ~ treatment_b  + (1|Atoll),
      prior = c(prior(normal(0,1), class = "b"), ##weakly informative, allow to vary, sd = 1 bc log link
                prior(normal(1,1), class = "Intercept")), ##weakly informative, mean = 2 bc log(mean(gw))
      iter = 3000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.9999), 
      sample_prior="yes",
      file = "popn_gonad_mod_hg_2b")
print(popn_gonad_mod_hg_2b) 
plot(popn_gonad_mod_hg_2b) 
pp_check(popn_gonad_mod_hg_2b) #looks weird - I think bc a lot of zeros in data now


#intercept:
posterior_samples(popn_gonad_mod_hg_2b) %>% 
  select(b_Intercept, prior_Intercept) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#slope:
posterior_samples(popn_gonad_mod_hg_2b) %>% 
  select(b_treatment_bNo_rats, prior_b) %>% 
  gather(Type, value) %>% 
    ggplot(aes(value, col=Type)) +
    geom_density()

#Compile posterior distributions:
popn_gonad_post <- posterior_samples(popn_gonad_mod_hg_2b)
median_hdi(popn_gonad_post$b_treatment_bNo_rats, .width = .95) #median estimate = 0.29, ymin = -0.72, ymax = 1.30


#test hypothesis that birdy effect is > 0
hypothesis(popn_gonad_mod_hg_2b, "treatment_bNo_rats>0") 
#evidence ratio = 2.8, posterior probability =  0.74
##so switches a bit to positive birdy effect, but still not much happening
```



